OptiFit: an improved method for fitting amplicon sequences to existing OTUs

### 2021-11-07 paper/paper.Rmd

close_enough <- function(x, y, tol = 10^-2) {
  assert_that(all(near(x, y, tol = tol)))
}
```

@@

```{r assert_db_mccs, include = FALSE}
mcc_decreasing <- c(closed_fit_rdp_mcc, closed_fit_silva_mcc, closed_fit_gg_mcc,
                    open_fit_silva_mcc, open_fit_gg_mcc, open_fit_rdp_mcc,
                    opticlust_mcc)
assert_that(!is.unsorted(mcc_decreasing))
```

@@

```{r assert_db_frac, include = FALSE}
assert_that(!is.unsorted(c(frac_fit_rdp, frac_fit_silva, frac_fit_gg)))
assert_that((frac_fit_gg - frac_fit_silva) < 10)
assert_that((frac_fit_gg - frac_fit_rdp) > 40)
```

@@

```{r assert_db_sec, include = FALSE}
assert_that(!is.unsorted(c(closed_fit_silva_human_sec,
                           opticlust_human_sec,
                           open_fit_silva_human_sec)))
```

@@

```{r assert_split_simple, include = FALSE}
close_enough(mcc_opticlust_vs_fit_split_simple, 0.02)
assert_that((frac_fit_split - frac_fit_gg) > 20)
assert_that(!is.unsorted(c(sec_closed_fit_split_vs_db, sec_closed_fit_db_vs_clust, sec_open_fit_split_vs_clust, sec_open_fit_split_vs_db)))
```

@@

```{r assert_split_strat, include = FALSE}
close_enough(c(mcc_fit_split_simple, mcc_fit_split_abun), mcc_fit_split_dist)
assert_that(!is.unsorted(c(sec_fit_split_dist,
                           sec_fit_split_simple,
                           sec_fit_split_abun)))
close_enough(frac_fit_split_simple, frac_fit_split_abun)
assert_that(frac_fit_split_dist < frac_fit_split_simple)
```

@@

```{r assert_ref_sizes, include = FALSE}
ref_seqs <- dn_dbs %>% pull(nseqs) 
names(ref_seqs) <- dn_dbs %>% pull(ref)
assert_that(!is.unsorted(ref_seqs[['rdp']], 
                         ref_seqs[['silva']], 
                         ref_seqs[['gg']])
            )
```

@@

We implemented the analysis workflow in Snakemake [@koster_snakemake_2012]
and wrote scripts in R [@r_core_team_r_2020],
Python [@van_rossum_python_2009],
and GNU bash [@noauthor_bash_nodate].
Software used includes mothur v1.46.1 [@schloss_introducing_2009],
VSEARCH v2.15.2 [@rognes_vsearch_2016],
the tidyverse metapackage [@wickham_welcome_2019],
R Markdown [@xie_r_2018],
ggraph [@pederson_ggraph_2021],
ggtext [@wilke_ggtext_2020],
numpy [@harris_array_2020],
the SRA toolkit [@noauthor_sra-tools_nodate],
and conda [@noauthor_anaconda_2016].
The complete workflow and supporting files required to reproduce this manuscript
are available at https://github.com/SchlossLab/Sovacool_OptiFit_2021.
<!-- TODO: update repo link with journal title -->

>>>>

# check whether all elements of two numeric vectors are close enough
close_enough <- function(x, y, tol = 10^-2) {
  assert_that(all(near(x, y, tol = tol)))
}

# check whether all elements given are sorted in non-descending order
is_nondesc <- function(...) {
  things <- c(...)
  if (length(things) < 1) {
    warning('Zero elements were given to `is_nondesc()`')
  }
  first <- things[1]
  last <- things[length(things)]
  !is.unsorted(things) & first <= last
}
```

@@

```{r assert_db_mccs, include = FALSE}
mcc_ascending <- c(closed_fit_rdp_mcc, closed_fit_silva_mcc, closed_fit_gg_mcc,
                    open_fit_silva_mcc, open_fit_gg_mcc, open_fit_rdp_mcc,
                    opticlust_mcc)
is_nondesc(mcc_ascending)
```

@@

```{r assert_db_frac, include = FALSE}
is_nondesc(frac_fit_rdp, frac_fit_silva, frac_fit_gg)
assert_that((frac_fit_gg - frac_fit_silva) < 10)
assert_that((frac_fit_gg - frac_fit_rdp) > 40)
```

@@

```{r assert_db_sec, include = FALSE}
is_nondesc(closed_fit_silva_human_sec,
           opticlust_human_sec,
           open_fit_silva_human_sec)
```

@@

```{r assert_split_simple, include = FALSE}
assert_that(mcc_opticlust_vs_fit_split_simple < 0.09)
assert_that((frac_fit_split - frac_fit_gg) > 20)
is_nondesc(sec_closed_fit_split_vs_db, sec_closed_fit_db_vs_clust, sec_open_fit_split_vs_clust, sec_open_fit_split_vs_db)
```

@@

```{r assert_split_strat, include = FALSE}
close_enough(c(mcc_fit_split_simple, mcc_fit_split_abun), mcc_fit_split_dist)
is_nondesc(sec_fit_split_dist,
           sec_fit_split_abun,
           sec_fit_split_simple)
is_nondesc(frac_fit_split_dist,
           frac_fit_split_abun,
           frac_fit_split_simple)
```

@@

```{r assert_ref_sizes, include = FALSE}
ref_seqs <- dn_dbs %>% pull(nseqs) 
names(ref_seqs) <- dn_dbs %>% pull(ref)
is_nondesc(ref_seqs[['rdp']],
           ref_seqs[['silva']],
           ref_seqs[['gg']])
```

@@

We implemented the analysis workflow in Snakemake [@koster_snakemake_2012]
and wrote scripts in R [@r_core_team_r_2020],
Python [@van_rossum_python_2009],
and GNU bash [@noauthor_bash_nodate].
Software used includes mothur v1.47.0 [@schloss_introducing_2009],
VSEARCH v2.15.2 [@rognes_vsearch_2016],
the tidyverse metapackage [@wickham_welcome_2019],
R Markdown [@xie_r_2018],
ggraph [@pederson_ggraph_2021],
ggtext [@wilke_ggtext_2020],
numpy [@harris_array_2020],
the SRA toolkit [@noauthor_sra-tools_nodate],
and conda [@noauthor_anaconda_2016].
The complete workflow and supporting files required to reproduce this manuscript
are available at https://github.com/SchlossLab/Sovacool_OptiFit_2021.
<!-- TODO: update repo link with journal title -->
